TacticToe: Learning the Game of Tactics

### 2018-02-08 tactictoe_jv.tex
graphic update and minor comments

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=0.5*\textwidth,
  height=0.5*\textwidth,
  ymin=0, ymax=3000,
  xmin=0, xmax=10,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
\addplot table[x=length, y=proofs] {data/original_proof_length};
\end{axis}
\end{tikzpicture}
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=0.5*\textwidth,
  height=0.5*\textwidth,
  ymin=0, ymax=120,
  xmin=10, xmax=50,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
\addplot table[x=length, y=proofs] {data/original_proof_length};
\legend{\tactictoe proofs,original proofs}
\end{axis}
\end{tikzpicture}
\caption{Number of original human proofs and (minimized see Section) \tactictoe
proofs with
a number $x$ of tactic
units. For easier comparison of the distributions, the number of \tactictoe
proofs was
scaled by the inverse of \tactictoe's success rate $\frac{1}{0.662}$.}
\end{figure}

There are more in proportion \tactictoe small proofs than original human proofs.

Increasing the timeout to more than 10 seconds would probably be beneficial as
we can see that the slope of the curve is far for being flat in Fig~..

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=20,
  ymin=0, ymax=100,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=oplen, y=solved] {data/tactictoe_by_oplen};
\addplot table[x=oplen, y=solved] {data/eprover_by_oplen};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{Percentage of problem solved with respect to the length of the
original proof until length 20.}
\end{figure}

%60 seconds experiment +
%adding theorems from the namespace and removing
%additionally saved theorems +
%recording proofs from Prove.

%
%The progress cannot be attributed only to the time increase but to testing and
%debugging of parameters.

\todoi{Remove feature vectors from the paper}
\section{Proof Recording}\label{sec:recording}

Recording proofs in an LCF-style proof assistant can be done at different
levels.
In \holfour all existing approaches relied on modifying the kernel. This was
used
either to export the primitive inference
steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable
for our
purpose of learning proving strategies at the intermediate tactic level. We
therefore
discuss recording proofs in an LCF-style proof assistant, with the focus on
\holfour
in this section.

Rather than relying on the underlying programming language, we parse the
original theory file containing proof scripts. This enables us to extract the
code of each tactic.
Working with the string representation of a tactic is better that working with
its value:
\begin{itemize}
\item Equality between two functions is easy to do from
their string representation. The direct consequence is that we can avoid
predicting the same tactic over and over again.
\item
The
results of
\tactictoe can be returned in a readable form to the user. In this way, the
user can learn from the feedback, analyze the proof and possibly improve on it.
An other point is that \tactictoe does not need to be installed to run
proof scripts generated by \tactictoe. In this way, the further development of
\tactictoe are not affecting the robustness of \holfour.
\item
It is difficult (probably impossible) to transfer \sml values
between theories as they are not run in the same \holfour session. In contrast,
\sml code can be exported, imported.
\item To produce a tactic value form a code is easy in \sml, which can be
achieved by using a reference to a tactic and update it with the function
\texttt{use}.
\end{itemize}

>>>>

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=0.5*\textwidth,
  height=0.5*\textwidth,
  ymin=0, ymax=3000,
  xmin=0, xmax=10,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
\addplot table[x=length, y=proofs] {data/original_proof_length};
\end{axis}
\end{tikzpicture}
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=0.5*\textwidth,
  height=0.5*\textwidth,
  ymin=0, ymax=120,
  xmin=10, xmax=50,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
\addplot table[x=length, y=proofs] {data/original_proof_length};
\legend{\tactictoe proofs,human proofs}
\end{axis}
\end{tikzpicture}
\caption{\label{fig:distrib}Number of proofs (y axis) with respect to their 
length (x axis) in 
number of tactic units. Comparison between the set of \tactictoe proofs and the 
set of human proofs.}
\end{figure}

In order to appreciate the difficulty of the evaluation set from a human 
perspective, the length distribution of human proofs in the 
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the 
graph that most of the proofs are short. The proofs found by tactictoe follows 
about the same 
distribution. This means that if \tactictoe finds a proof, there is about a 50 
percent chance that it will be shorter than what a human would come up with.
In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to 
see how well \tactictoe and \eprover performs on increased level of 
difficulties. As expected, the longer the original proof is, the harder it is 
for \tactictoe and \eprover to re-prove the theorem by their own mean. 
The performance of \tactictoe is compelling with more than half of the theorem 
that required a proof of length six being reproven. It is also consistently 
better than \eprover for any proof length.

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=20,
  ymin=0, ymax=100,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=oplen, y=solved] {data/tactictoe_by_oplen};
\addplot table[x=oplen, y=solved] {data/eprover_by_oplen};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{\label{fig:percentage}Percentage of problem solved with respect to the 
length of the
original proof until length 20.}
\end{figure}

\section{Proof Recording}\label{sec:recording}
Recording proofs in an LCF-style proof assistant can be done at different
levels.
In \holfour all existing approaches relied on modifying the kernel. This was
used
either to export the primitive inference
steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable
for our
purpose of learning proving strategies at the intermediate tactic level. We
therefore
discuss recording proofs in an LCF-style proof assistant, with the focus on
\holfour
in this section.
<
Rather than relying on the underlying programming language, we parse the
original theory file containing proof scripts. This enables us to extract the
code of each tactic.
Working with the string representation of a tactic is better that working with
its value:
\begin{itemize}
\item Equality between two functions is easy to do from
their string representation. The direct consequence is that we can avoid
predicting the same tactic over and over again.
\item
The
results of
\tactictoe can be returned in a readable form to the user. In this way, the
user can learn from the feedback, analyze the proof and possibly improve on it.
An other point is that \tactictoe does not need to be installed to run
proof scripts generated by \tactictoe. In this way, the further development of
\tactictoe are not affecting the robustness of \holfour.
\item
It is difficult (probably impossible) to transfer \sml values
between theories as they are not run in the same \holfour session. In contrast,
\sml code can be exported, imported.
\item To produce a tactic value form a code is easy in \sml, which can be
achieved by using a reference to a tactic and update it with the function
\texttt{use}.
\end{itemize}
